---
title: "Homework_4"
author: Ningyu Han
output: github_document
date: "2023-02-27"
---

```{r}
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'
library('manipulate')
library('splines') ## 'ns'
library('caret') ## 'knnreg' and 'createFolds'
```

```{r}
y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
```

## Randomly split the mcycle data into training (75%) and validation (25%) subsets.
```{r}
dat <- mcycle
set.seed(123)

n <- nrow(dat)
train_indices <- sample(1:n, round(0.75*n), replace = FALSE)
mcycle_train <- mcycle[train_indices, ]
mcycle_test <- mcycle[-train_indices, ]
```

```{r}
y_valid <- mcycle_test$accel
x_valid <- matrix(mcycle_test$times, length(mcycle_test$times), 1)

plot(x_valid, y_valid, xlab="Time (ms)", ylab="Acceleration (g)")
```

## Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values. (hint: the script already implements this)
```{r}
## Epanechnikov kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## lambda - bandwidth (neighborhood size)
kernel_epanechnikov <- function(x, x0, lambda=1) {
  d <- function(t)
    ifelse(t <= 1, 3/4*(1-t^2), 0)
  z <- t(t(x) - x0)
  d(sqrt(rowSums(z*z))/lambda)
}
```

```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, knn=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:knn]] <- 1
  
  return(w)
}
```

```{r}
## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}

yhat <- nadaraya_watson(y, x, x,kernel_k_nearest_neighbors, knn=1)
```

```{r}
## Helper function to view kernel (smoother) matrix
matrix_image <- function(x) {
  rot <- function(x) t(apply(x, 2, rev))
  cls <- rev(gray.colors(20, end=1))
  image(rot(x), col=cls, axes=FALSE)
  xlb <- pretty(1:ncol(x))
  xat <- (xlb-0.5)/ncol(x)
  ylb <- pretty(1:nrow(x))
  yat <- (ylb-0.5)/nrow(x)
  axis(3, at=xat, labels=xlb)
  axis(2, at=yat, labels=ylb)
  mtext('Rows', 2, 3)
  mtext('Columns', 3, 3)
}

matrix_image(attr(yhat, 'k'))
```

## With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.
```{r}
## Compute effective df using NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## kern  - kernel function to use
## ... - arguments to pass to kernel function
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}
```

```{r}
x_train = matrix(mcycle_train$times, length(mcycle_train$times), 1)
yhat_train = nadaraya_watson(mcycle_train$accel, x_train, x_train, kernel_k_nearest_neighbors, knn=1)
```

```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
  
loss_squared_error <- function(y, yhat)
  (y - yhat)^2

## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

train_error <- rep(NA, 50)
for (i in 1:50){
  yhat_train <- nadaraya_watson(mcycle_train$accel, x_train, x_train, kernel_k_nearest_neighbors, knn = i)
  train_error[i] <- error(mcycle_train$accel, yhat_train, loss = loss_squared_error)
}

plot(seq(1:50),train_error, xlab = "k", ylab = "training error")
```

```{r}
y = mcycle_train$times
x <- x_train

## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom

bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d

## make predictions using NW method at training inputs
yhat <- nadaraya_watson(y, x, x,
  kernel_epanechnikov, lambda=5)

## view kernel (smoother) matrix
matrix_image(attr(yhat, 'k'))

## compute effective degrees of freedom
edf <- effective_df(y, x, kernel_epanechnikov, lambda=5)
aic(y, yhat, edf)
bic(y, yhat, edf)
```

```{r}
aic_error <- rep(NA, 50)
for(i in 1:50){
  yhat_train <- nadaraya_watson(mcycle_train$accel, x_train, x_train,kernel_k_nearest_neighbors, knn=i)
  edf <- effective_df(y, x, kernel_k_nearest_neighbors, knn=i)
  aic_error[i] <-  aic(y, yhat_train, edf)
}
print(aic_error)
plot(seq(1:50), aic_error, xlab = 'k',ylab = 'AIC')
```

```{r}
bic_error <- rep(NA, 50)
for(i in 1:50){
  yhat_train <- nadaraya_watson(mcycle_train$accel, x_train, x_train,kernel_k_nearest_neighbors, knn=i)
  edf <- effective_df(y, x, kernel_k_nearest_neighbors, knn=i)
  bic_error[i] <- bic(y, yhat_train, edf)
}
print(bic_error)
plot(seq(1:50), bic_error, xlab = 'k',ylab = 'BIC')
```

```{r}
valid_error <- rep(NA, 50)
for(i in 1:50){
  yhat_valid <- nadaraya_watson(mcycle_test$accel, x_valid, x_valid,kernel_k_nearest_neighbors, knn=i)
  valid_error[i] <-  error(mcycle_test$accel,yhat_valid)
}

#Error
#training error
plot(seq(1:50), valid_error, xlab = 'k', ylab = 'testing error')
```

```{r}
## create a grid of inputs 
x_plot <- matrix(seq(min(x),max(x),length.out=100),100,1)

## make predictions using NW method at each of grid points
y_hat_plot <- nadaraya_watson(y, x, x_plot,
  kernel_epanechnikov, lambda=1)

## plot predictions
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
```

```{r}
# # how does k affect shape of predictor and eff. df using k-nn kernel ?
# manipulate({
#   ## make predictions using NW method at training inputs
#   y_hat <- nadaraya_watson(y, x, x,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   edf <- effective_df(y, x,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   aic_ <- aic(y, y_hat, edf)
#   bic_ <- bic(y, y_hat, edf)
#   y_hat_plot <- nadaraya_watson(y, x, x_plot,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
#   legend('topright', legend = c(
#     paste0('eff. df = ', round(edf,1)),
#     paste0('aic = ', round(aic_, 1)),
#     paste0('bic = ', round(bic_, 1))),
#     bty='n')
#   lines(x_plot, y_hat_plot, col="#882255", lwd=2)
# }, k_slider=slider(1, 10, initial=3, step=1))
```

## For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.
```{r}
## 5-fold cross-validation of knnreg model
## create five folds
set.seed(1985)
mcycle_flds  <- createFolds(mcycle$times, k=5)
print(mcycle_flds)

sapply(mcycle_flds, length)  ## not all the same length

cvknnreg <- function(kNN = 10, flds=mcycle_flds) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    mcycle_trn <- mcycle[-flds[[tst_idx]],]
    mcycle_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel ~ times,
                      k=kNN, data=mcycle_trn)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, mcycle_tst)
    cverr[tst_idx] <- mean((mcycle_tst$accel - pre_tst)^2)
  }
  return(cverr)
}

## Compute 5-fold CV for kNN = 1:50
cverrs <- sapply(1:50, cvknnreg)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:50)
cverrs_mean <- apply(cverrs, 2, mean)
cverrs_sd   <- apply(cverrs, 2, sd)
```

## Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).
```{r}
## Plot the results of 5-fold CV for kNN = 1:50
plot(x=1:50, y=cverrs_mean, 
     ylim=range(cverrs),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:50, x1=1:50,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=20)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3)
```

## Interpret the resulting figures and select a suitable value for the tuning parameter.
**As we can see in the plot above, when k = 6, the model has the lowest test error. However, when you look at the cverrs_mean and cverrs_sd, we find that, when k = 4, it has the lowest standard deviation 134.9745, its test error is also low. As a result, I think k = 4 may be a suitable value for the tuning parameter.**




